---
title: "Lobster Exploration"
author: "Felicia Cruz"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in the data 

- Use `here()` to read in the data from the docs folder
- NAs are -99999 in the original dataset 
- Use `clean_names()` to convert the column names to lower snake case
```{r}
lobster_abundance <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>%
                                clean_names()

```

- Use `uncount()` to convert the data from frequency to case format on the existing `count` column
```{r}
lobster_abundance <- lobster_abundance %>% 
  uncount(count)
```
`dplyr::uncount()` makes each observation its own row. For example, if the `count` column showed that there were 3 lobsters in a transect all with the same size, this will make it so there are now three separate rows representing those three observations. 

## Find counts and mean sizes by site and year 

```{r}
lobster_abundance_counts_and_means <- lobster_abundance %>% 
  group_by(site, year) %>% 
  summarize(mean_length = mean(size_mm, na.rm = TRUE),
            total_count = n())
```
```{r}
lobster_carapace_lengths <- lobster_abundance_counts_and_means %>% 
  ggplot(aes(x = year, y = total_count)) +
  geom_point(aes(color = site)) 

lobster_carapace_lengths
```

## Proportion of legal lobsters at each site for 2020

```{r}
subset_2020 <- lobster_abundance %>%
  filter(year == 2020) 

subset_2020_summary <- subset_2020 %>% 
  mutate(legal = case_when(
    size_mm >= 79.76 ~ "yes",
    size_mm < 79.76 ~ "no"
  )) %>%
  group_by(site, legal) %>% 
  summarize(count = n())
  
stacked_plot <- subset_2020_summary %>% 
  ggplot(aes(x = site, y = count)) +
  geom_col(position = "fill", aes(fill = legal))

stacked_plot
```
Isla Vista and Naples are the two sites with the largest proportion of legal lobsters in 2020. Both of these sites are in marine protected areas. 


### Random lobster wrangling 

`filter()` practice
```{r}
lobsters <- lobster_abundance

ex_a <- lobsters %>% 
  filter(site == "IVEE" | site == "CARP" | site == "NAPL")

ex_b <- lobsters %>% 
  filter(month == 8)

ex_c <- lobsters %>%  
  filter(site == "AQUE" | size_mm > 70)

ex_d <- lobsters %>%  
  filter(site != "NAPL")
```

`group_by() %>% summarize()` practice
```{r}
ex_e <- lobsters %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE))
ex_f <- lobsters %>% 
  group_by(site, month) %>% 
  summarize(max_length = max(size_mm, na.rm = TRUE))

```

`mutate()` practice
```{r}
ex_g <- lobsters %>% 
  mutate(size_cm = size_mm / 10)

ex_h <- lobsters %>% 
  mutate(site = tolower(site))

ex_i <- lobsters %>% 
  mutate(area = as.character(area))
```

`case_when()` practice 
```{r}
ex_j <- lobsters %>% 
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small",
    size_mm > 70 ~ "large"
  ))

ex_k <- lobsters %>% 
  mutate(designation = case_when(
    site == "IVEE" | site == "NAPL" ~ "MPA",
    TRUE ~ "not MPA"
  ))
```

